Generalized Plackett-Luce Likelihoods

### 2022-08-10 inst/jss_hyper3.Rnw
abstract

<<setup, cache=FALSE,echo=FALSE>>=
library("knitr")
opts_chunk$set(cache=TRUE, autodep=TRUE)
@

@@

show how the \proglang{map} class of the Standard Template Library are
used: \proglang{weightedplayervector} objects map strings to long
doubles (specifically, mapping player names to their multiplicities),
and objects of class \proglang{hyper3} are maps from a
\proglang{weightedplayervector} object to long doubles.  One advantage
of this is efficiency: Search, removal, and insertion operations have
logarithmic complexity.  As an example, the following \proglang{C++}
pseudo code would create a log-likelihood function for the first term
in Equation~\ref{plackettluce}:

@@

<<>>=
library("hyper2")
H <- hyper3()
H[c(a=1)] <- 1
H[c(a=3,b=2,c=1)] <- -1
H
@

@@

\bibliography{chess}

\end{document}

>>>>

<<setup, cache=FALSE,echo=FALSE>>=
library("knitr")
library("hyper2")
opts_chunk$set(cache=TRUE, autodep=TRUE)
@

@@

show how the \proglang{map} class of the Standard Template Library is
used: \proglang{weightedplayervector} objects map strings to long
doubles (specifically, mapping player names to their multiplicities),
and objects of class \proglang{hyper3} are maps from a
\proglang{weightedplayervector} object to long doubles.  One advantage
of this is efficiency: Search, removal, and insertion operations have
logarithmic complexity.  As an example, the following \proglang{C++}
pseudo code would create a log-likelihood function for the first term
in Equation~\ref{plackettluce}:

@@

<<>>=
H <- hyper3()
H[c(a=1)] <- 1
H[c(a=3,b=2,c=1)] <- -1
H
@

@@

\section{Conclusions and further work}

Plackett-Luce likelihood functions for rank datasets have been
generalized to impose identity of Bradley-Terry strengths for certain
groups; the preferred interpretation is a running race in which the
competitors are split into equivalence classes of clones.
Implementing this in \proglang{R} is accomplished via a \proglang{C++}
back-end making use of the \proglang{STL} ``map" class which offers
efficiency advantages, especially for large objects.

New likelihood functions for simple cases with three or four
competitors were presented, and extending to larger numbers furnishes
a generalization of the Mann-Whitney-Wilcoxon test that offers a
specific alternative (Bradley-Terry strength) with a clear operational
definition.  Further generalizations allow the analysis of more than
two groups, here applied to Olympic Javelin throw distances.
Generalized Plackett-Luce likelihood functions were used to assess the
constructors' championship and a reasonable null.  Specifically, the
hypothesis that the strength of the Mercedes team remained unchanged
between 2020 and 2021 was tested and rejected.

Further work might include a systematic comparison between
\code{hyper3} approach and the Mann-Whitney-Wilcoxon test, including
the characterisation of the power function of both tests.  Further,
the package could easily be extended to allow non-integer
multiplicities, which might prove useful in the context of reified
Bradley Terry techniques~\citep{hankin2020}.

\bibliography{chess}
\end{document}
